ROC curve analysis

### 2021-09-30 Reports/FFCleaning2.Rmd
September 30 updated pipeline and results

ggplot(cellbytime, aes(x=childteen, y=Proportion, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+facet_wrap(~Celltype, labeller=labeller(Celltype=celltype_labels))+stat_compare_means(label="..p.signif..")+ggtitle("Cell type proportions by visit")+theme_classic()+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()
```
```{r celltypes by visit spaghetti}
cairo_ps(file.path(outputdir, 'SpaghettiCellTypes.ps'))
ggplot(cellbytime, aes(x=childteen, y=Proportion, group=idnum))+geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~Celltype, labeller=labeller(Celltype=celltype_labels))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))+ylab('Proportion')+theme(text=element_text(size=16))
dev.off()
```

>>>>

ggplot(cellbytime, aes(x=childteen, y=Proportion, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+facet_wrap(~Celltype, labeller=labeller(Celltype=celltype_labels))+stat_compare_means(label="..p.signif..")+ggtitle("Cell type proportions by visit")+theme_classic()+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()
```

```{r celltypes by visit spaghetti}
cairo_ps(file.path(outputdir, 'SpaghettiCellTypes.ps'))
ggplot(cellbytime, aes(x=childteen, y=Proportion, group=idnum))+geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~Celltype, labeller=labeller(Celltype=celltype_labels))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))+ylab('Proportion')+theme(text=element_text(size=16))
dev.off()
```

### 2021-09-30 Reports/ROC_report.Rmd
September 30 updated pipeline and results

global_roc%>%dplyr::select(covariates, methylation, childteen, test)%>%unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric'))%>%
  ggplot(aes(x=spec, y=sens, color=covariates))+geom_line()+
  facet_grid(childteen~methylation, labeller=labeller(childteen=age_labels, methylation=outcome_labels))+
  theme_classic()+theme(legend.position = 'bottom')+guides(color=guide_legend(nrow=6,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')

#is prediction better when using PMS? 
global_roc%>%dplyr::select(predictors, covariates, methylation, childteen, test)%>%
  mutate(modeltype=ifelse(str_detect(predictors, '\\+'), covariates, 'No covariates'))%>%
  unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') &
           !str_detect(covariates, 'm1g2|sv'))%>%
  ggplot(aes(x=spec, y=sens, color=methylation, linetype=modeltype))+geom_line()+
  facet_grid(childteen~., labeller=labeller(childteen=age_labels))+
  scale_color_discrete(labels=outcome_labels)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')

@@

#is best methylation predictor consistent across ancestries
local_roc %>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'SSnewbornCT|cg055|global|pediatric') & covariates=='')%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()

#things get weird when using the base model vars
local_roc %>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'SSnewbornCT|cg055|global|pediatric') & covariates!='')%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()

#investigated change in ROC when adding in each base model var individually
local_roc_test%>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'global'))%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()
#big jump in ROC curve for European & hispanic - batch variable 
#big jump in ROC curve for European - poverty and hispanic - local pcs
modeldata %>%filter(ancestry=='Hispanic ancestry' & childteen=='T')%>%xtabs(~Sample_Plate+smkPreg_binary, data=.)

#is prediction improved when using two time points?

>>>>

global_roc%>%dplyr::select(covariates, methylation, childteen, test)%>%unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric'))%>%
  ggplot(aes(x=spec, y=sens, color=covariates))+geom_line()+
  facet_grid(childteen~methylation, labeller=labeller(childteen=age_labels, methylation=outcome_labels))+
  theme_classic()+theme(legend.position = 'bottom')+guides(color=guide_legend(nrow=6,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')
```

```{r}

#is prediction better when using PMS? 
global_roc%>%dplyr::select(predictors, covariates, methylation, childteen, test)%>%
  mutate(modeltype=ifelse(str_detect(predictors, '\\+'), covariates, 'No covariates'))%>%
  unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') &
           !str_detect(covariates, 'm1g2|sv'))%>%
  ggplot(aes(x=spec, y=sens, color=methylation, linetype=modeltype))+geom_line()+
  facet_grid(childteen~., labeller=labeller(childteen=age_labels))+
  scale_color_discrete(labels=outcome_labels)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')
```

```{r}

@@

```{r}
#is best methylation predictor consistent across ancestries
local_roc %>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'SSnewbornCT|cg055|global|pediatric') & covariates=='')%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()

#things get weird when using the base model vars
local_roc %>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'SSnewbornCT|cg055|global|pediatric') & covariates!='')%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()
```

```{r}
#investigated change in ROC when adding in each base model var individually
local_roc_test%>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'global'))%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()
#big jump in ROC curve for European & hispanic - batch variable 
#big jump in ROC curve for European - poverty and hispanic - local pcs
modeldata %>%filter(ancestry=='Hispanic ancestry' & childteen=='T')%>%xtabs(~Sample_Plate+smkPreg_binary, data=.)

```{r}
#is prediction improved when using two time points?
